
While this book is obstensibly about how \emph{computers} answer
questions, that is not where the book begins.
%% After a brief
%% introduction laying out the organization of the book, we go to the
%% \emph{history} of humans asking each other questions.
%
This history is important because it explains why humans value
answering questions.
%
This provides historical context on why humans equate answering
questions correctly with intelligence but also because key elements for
artificial intelligence---item response theory and preference models---were
developed from looking at how humans answered questions.

The next stop is looking at \emph{today's} question answering,
begining with Alan Turing, arguably the father of computing and
artificial intelligence.
%
Two wartime efforts---one at air and one at sea---began a rivalry in
question answering, which we call the Cranfield and Manchester based
on the college towns home to Alan Turing's theorizing on intelligent
computing and Cyril Cleverdon's practical evaluation of computing
systems.
%
Having set the stage of this rivalry, we then see how these ideas
built the foundations of modern \abr{ai} from the methods that let
\watson{} win on \jeopardy{}
%
\ifproposal
(this is the subject of the sample chapter for the proposal)
\fi
%
to datasets like Google's Natural Questions that set the stage for
smart chatbots and the models like \abr{gpt}.

Finally, the book closes with the \emph{future} of question answering,
focusing on what it means for an \abr{ai} to be intelligent: how to
turn Alan Turing's thought experiment into reality and how to make
sure that computers don't trick us into thinking they're intelligent
(but just cheating on the test).
%
And if computers are becoming intelligent the question is how to
\emph{align} them with users, which sees the return of item response
theory and preference models: this time built up so we can better
understand why \abr{ai} says the things it does.
%
Finally, we close with a discussion of the ``big questions'' that
computers cannot answer on their own: how can the models developed
from answering questions from humans best work alongside humans for
work and play.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:
