

With many tasks in computer science, asking the question ``where did this task come from'' has a simple answer usually in the form of an acronym . 
%
Character recognition was sparked by the \abr{mnist} dataset; entailment was sparked by \abr{snli}; sentiment classification was defined by \citet{}.

And while there are particular \emph{flavors} of question answering that habe been defined to make life easier or more discriminative for computers, the basic task of question in, answer out is natural to any inquisitive three year old.
%
Of course, question answering is not the only task like this: translation likewise is recognizable to any stranger in a strange land trying to cope with an unfamiliar language: dozens of books have been written on that subject, my favorite is \textit{Le Ton Beau de Marot}.

Question answering's roots likewise run deep.  In this chapter, I want to argue that question answering is not just a useful task but one that is woven deep into our Jungian subconsious.
%
Answering questions can define your identity, reveal fundamental truths of the universe, and create a shared reality.
%
This chapter highlights these pillars of question answering and how these questions of identity and knowledge shape not just our modern world but a world where our reality is also partially defined by artificial intelligence.

\section{The Sphynx}



In Greek myth, the Sphynx asked everyone who entered the city a riddle.  Here’s PDQ Bach’s interpretation of the riddle:
\begin{quote}
What starts out on four legs then goes round on two
Then finishes on three before it’s through
\end{quote}
And the answer is:
\begin{quote}
Well a baby then a man, that’s plain, 
then finally an old geezer on a cane
\end{quote}

Now I won’t spoil the whole story, but by answering the question correctly, Oedipus revealed that he was smarter than the average bear and also unlocked secrets from his past.  In a commencement address to Whitman college, Dana Burgess explained the difference between this and Google’s Natural questions and how that relates to the mission of professors (like me):

\begin{quote}
Nobody told Oedipus who he was; he figured that out for himself.  ... You can do a Google search to find out the capital of Arkansas (Little Rock?) but you can’t do a Google search to find out who you are, what you’re good at, what makes you happy, what matters for your life.  Information transferal isn’t any help with that stuff.
\end{quote}

When folks say that a college professor is like a sphinx, they don’t usually mean that in a nice way.  It suggests that we know the answers students need, but we keep them secret for the sake of our own power.  … That’s part of the riddle of [The Sphynx’s] hybrid nature: the curse of being forced to solve a riddle is also the gift of the ability to solve riddles. 

\section{Gestumblindi}

In Norse myth, Gestumblindi got into some trouble with King Heidrek and asked the Gods for help.
%
In response, Odin came to Gestumblindi---looking exactly like him---and stood in his place in front of Heidrek.
%
What took place was an interrogation that went in two directions: King Heidrek attempted to determine who this Gestumblindi-looking person was in front of him and Odin attempted to see how smart King Heidrek was.

What I like about this story is that it presages many of the themes that we will see with \abr{ai}: ordinary people thinking that \abr{ai} will help them, an interrogation contest to see who is human or not, and an impersonation.
%
It also presages many of our fears about \abr{ai}: Odin's final question to Heidrek is:
\begin{quote}
what had he whispered (or will whisper) to his son, Baldur, before his son is to
be taken to his funeral pyre.
\end{quote}
We talk in this book frequently about bad questions, but this a particularly bad question.  Not only is it something that only Odin knows, it probably has not happened yet (depending on what you believe about the circularity of time), and when it does happen it will end the world by bringing about Ragnar\"ok.  

% TODO(jbg): add forward pointers

Once Heidrek hears this question, he knows that the figure before him is not Gestumblindi but rather Odin.
%
And like a frustrated customer who thought they were talking to a human for ten minutes only to figure out that they're talking to a bot, Heidrek takes out his sword and attacks, only for Odin to turn into a bird and fly away.

\section{The Socratic Method}

Because I’m a professor, I am somewhat biased in my appreciation of questions: both in the classrooms and on exams.  But I hope one thing that I can convince you of is that these are important as foundations of civilization: the world would be a much worse place without these question answering activities.


Let’s begin with the socratic method.  If you haven’t heard of this before, you might have experienced it (especially if you’re in my classes).  Let’s hear how John Houseman explain what it is in the 1971 film The Paper Chase:
\begin{quote}
We use the Socratic Method here.  I call on you, ask you a question\dots and you answer it.  Why don't I just give you a lecture?  Because through my questions, you learn to teach yourselves.

Through this method of questioning, answering\dots questioning, answering\dots we seek to develop in you
the ability to analyze that vast complex of facts that constitute the relationships of members within a given society.  Questioning and answering.

At times you may feel that you have found the correct answer. I assure you that this is a total delusion on your part. You will never find the correct, absolute, and final answer.  In my classroom, there is always another question\dots another question to follow your answer.

Yes, you're on a treadmill.  My little questions spin the tumblers of your mind. You're on an operating table. My little questions are the fingers probing your brain.

We do brain surgery here.  You teach yourselves the law but I train your mind.  You come in here with a skull full of mush\dots and you leave thinking like a lawyer.
\end{quote}

I want to particularly hone in on the assertion that it ``trains your brain''.  While not quite mythological, the term ``Socratic Dialog'' goes back to ancient Greece.  Plato featured Socrates in his dialogues Euthyphro and Ion, giving the perspective of the student the student, the person answering the questions, struggling with multiple interpretations and ``showing their work'' as they emerge on the other side, enlightened by the exchange.

And indeed, one of the best moments for me as a professor is to see the aha moment in students eyes after they answer a question correctly.  And this is still how we train lawyers, who have to use argument and discussion to keep our society ordered!  But as we will see in the next chapter, that is not the only way the questions and answers create order in our civilization.



% - Gestumblindi: the art of making a riddle is prized, and it is the
% Gods who are best at the task

% Perhaps add snow white and Turing's fascination with it? 